### 2023-07-11 manuscript_english.tex
expand introduction

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Suitability prevalence area index in late quaternary explains genetic diversity in Tassel eared Squirrels},
  pdfauthor={Norma Hernandez, Angel Robles Nathan Upham},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Suitability prevalence area index in late quaternary explains
genetic diversity in Tassel eared Squirrels}
\author{Norma Hernandez, Angel Robles Nathan Upham}
\date{2023-06-27}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

The current distributions of species do not always correspond to their
historical distributions over evolutionarily significant periods. This
is because environmental conditions are not static over time; species
tend to distribute where conditions are most favorable.

It follows from this relationship between environment and species
distribution that population size also varies with time. This variation
in population size is related to the effective population size \(N_e\)
(Karlin 1968), so that an index reflecting changes in environmental
conditions in geography might be expected to be related to \(N_e\) and
thus to indicators of population structure such as the fixation index
\(F_{st}\). Thus, it is possible to relate patterns of changes in the
distribution of species to the genetic structure of their populations
using a statistical model that explains this relationship.

With this approach we can predict the geographic pattern of population
structure from environmental information. This approach is strongly
driven by advances in currently available climate simulations (Leonardi
et al. 2023; Krapp et al. 2021), as well as next generation sequencing
data, and supported by both ecological niche (Thorup et al. 2021;
Nogués-Bravo 2009) and population genetics theories (Lira-Noriega and
Manthey 2014).

In this work we propose a method to find the Suitability Prevalence Area
(SPA) as an index with a double purpose: 1) to find endemic areas to
delimit the historical distribution of the species and 2) to explain the
patterns of genetic diversity.

To obtain the SPA, we performed a historical reconstruction of the
geographical range back to 120 000 BC at 2 000 year intervals and
recorded the environmental suitability at each site.

Subsequently, we delimited historical endemic areas to locations where
the prevalence of suitability remained at 90\% during this period. From
the fixation index calculated with respect to populations in the
historical endemic areas, a statistical model of the fixation index as a
function of SPA was performed. With this statistical model, the fixation
index values were projected to the current distribution to obtain a map
with the geographic pattern of this index.

As a case study we consider squirrels (\emph{Sciurus aberti}), a species
currently distributed in disjunct patches from the southern Rocky
Mountains in the United States to the northern Sierra Madre Occidental
in Mexico for which it is possible to find reliable information on both
the genetic structure of its populations and its current distribution
(Bono et al. 2018; Burgin et al. 2018).

Our results reveal that suitability prevalence corresponds to the
fixation index of \emph{S. aberti} populations with respect to a source
population. Populations closer to the historical endemic area present a
higher genetic diversity and a lower \(F_{st}\) value. Finally, this
study allows us to add a biogeographic explanation to the results
obtained with population genetic methods and to generate maps of this
structure as tools to support conservation with a perspective that
integrates both population genetics and historical patterns of species
distribution.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Relating the genetic structure of the populations of a species to its
geographic distribution is a complicated challenge from the point of
view of biogeography. The task is made easier if evolutionary history is
incorporated on both sides of the relationship. On the one hand, giving
depth and an evolutionary perspective to the geographic distribution of
a taxon relies primarily on ecological niche theory, and is driven by
the massive amount of information currently available from both
digitized biological collections and citizen science (through which
records of taxa occurrence are through which records of taxa occurrence
are available) as well as from environmental information available for
geographic information systems and climate models. information systems
and accurate climatological models of both the present and the past.

This change in reproductive success is driven by a set of factors,
however we propose in this work a conceptual framework where we consider
only environmental factors at a macroecological scale, inspired by the
concept of Grinnellian niches (those composed by sets of points of
non-interactive variables in multidimensional spaces), seeking as a
principle the necessary conditions for fitness and not the response of
taxa. points of non-interactive variables in multidimensional spaces),
seeking as a principle the conditions necessary for fitness and not the
response of taxa.

That is, we simplify our model (as is done in the physical sciences) to
link through fitness the changes in biological populations and their
structure, observed through population genetics, with patterns of
spatial distribution over time. Thus, the goal is to link biogeographic
patterns and evolutionary processes of populations in a single
conceptual framework.

It is from this Grinellian niche where we start, assuming that taxa in
general have to be distributed in the geography where environmental
conditions are more favorable, noting that the distribution of taxa is
not static, depending on dynamic environmental conditions and therefore
the current known distribution of a taxon does not always correspond to
its historical distribution over time, since the current climate has not
remained so for periods of time on scales relative to the processes of
population genetics. (i. e. tens or hundreds of thousands of years).
Even for some species these dynamics seem to induce ecological processes
such as seasonal migrations in birds, or hibernation in mammals and
frogs. But in general, environmental changes are reflected in geography
as range expansions and contractions over time. That is, processes of
population dynamics, driven by environmental fitness, in our approach.

This change in population dynamics is implied by the genetic structure
of populations through the concept of effective population size \(N_e\)
(Wright 1931; Karlin 1968). So if we set a similar time scale for these
changes, we will be able to observe the fitness relationship (through
population dynamics) in both genetic structure and distribution range
changes over time. That is, we need an index that reflects environmental
changes in the geographic distribution of species and changes in the
genetic structure of the population. In this case, we can consider the
fixation index \(F_{st}\) as an indicator of the difference between
populations that come from a source population and that have undergone
range shifts that we assume are driven by environmental changes. Under
these assumptions we expect that it is possible to relate changes in the
distribution of taxa to the genetic structure of their populations using
statistical models to explain the relationship.

\hypertarget{theoretical-basis}{%
\subsubsection{Theoretical basis}\label{theoretical-basis}}

\hypertarget{environment-fitness-and-suitability}{%
\paragraph{Environment, fitness and
suitability}\label{environment-fitness-and-suitability}}

For a given region \(\mathbf{G}\) in te geography, at any point in time,
we have a set of existing environmental conditions
\(\mathbf{E}(\mathbf{G}, t) = \Big \{ \vec{e}_i; \mathbf{G}, t\Big \}\).
We define \(\vec{e}_i =(e_1, e_2, ..., e_v )_i\) as a vector of
environmental variables, that is, a point in the space of \(v\)
environmental variables \(\mathbf{E}(\mathbf{G}, t)\). Then
\(R_0(\vec{e}_i)\) is a function that relates fitness to environment,
that is, for a species with non-overlapping generations, the
population's net reproductive rate \(R_0\) is a measure of fitness that
depends on environmental conditions. If we take the set of places in
environmental space where these values are greater than a threshold, we
can define a volume in multidimensional space such that \(R(\vec{e})\)
is always positive and greater than this threshold, that is
\(\mathbf{N}_F=\{\vec{e} | R_0(\vec{e}) \ge k \}\). i. e. the set of
environmental conditions for which values of fitness are capable of
sustaining populations. We call
\(\mathbf{N}_F \subseteq \mathbf{E}(\mathbf{G}, t)\). This is a way of
expressing the fundamental niche as a function of environmental
conditions. So we can correspong each point of geography to
environmental space, with different environmental conditions and
different growth rates. However all sites within the region defined by
\(\mathbf{N}_F\) will have viable populations and positive rates
depending on time because we assume that the environment is dynamic and
each geographic point is static, so for each site in geographic space
this growth rate depends on the environment and therefore is different
at each time: \(R_0 = R_0(\vec{e}(t)) \}\). Thus locally the growth rate
will be related to the local population size as long as it is within the
boundary defined by \(\mathbf{N}_F\).

This equation was proposed by Soberon and Peterson Soberón and Peterson
(2020) to relate demographics to a function of fitness and environment.
In that study they tested the validity with monthly variables and
empirical values on a short time scale. relating life history parameters
and the environmental suitability of niche models.

This set can be seen as a convex volume in environmental space and we
relate this growth rate to an environmental suitability function bounded
between \(0\) and \(1\), for example, for a unimodal multivariate pdf,
it would be equivalent to the distance from the maximum of the function.
That is, the fitness-related environmental suitability can be any
function such that it is \(0\) if \(R_0(\vec{e}_i) > 1\) and \(1\) when
\(R_0\) is evaluated at the expected value of \(\vec {e}(t)\) for all
values of the set \(\mathbf{N}_F\) where \(R_0(\vec{e}_i) > 1\). In a
practical way, instead of \(\mathbf{N}_F\), we approximated
\(\mathbf{N}_F^{*}(\mathbf{G, t}) = \mathbf{E} \cap \mathbf{N}_F\) (the
existing niche) from known presence points in the geography
\(\mathbf{G}\) obtained from emprical observations stored in database,
and map to \(\mathbf{E}\) in order to numerically fitting a convex
volume of the environmental space that enclose the known information.

The objective of our approach is not to know the function of \(R_0\) but
rather to relate this local variation and the function \(\mathbf{N}_F\)
with a function \(S_i(\vec{e}_i)\) of environmental suitability that can
be estimated from known methods of applied mathematics and machine
learning (for example the case of Maxent, Phillips, Anderson, and
Schapire (2006)).

Population size locally is a function of environmental suitability,
where if suitability is \(0\) the population is \(0\) and if suitability
is \(1\) the population has a maximum growth rate \(R_i(\vec{e}_i)\) \[
N_i \propto  S_i(\vec{e}_i)
\]

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{all_figures/figure_1.png}
\caption{Hypothetical niche in environmental space. An arbitrary point
moves within the environmental space entering and exiting the volume
(niche) where the intrinsic growth rate is positive.}
\end{figure}

Once we have an approximation of this volume, we consider that each
point in the geography is static and has a trajectory in the
environmental space and for each trajectory the fitness function changes
and therefore the suitability function changes but not in a homogeneous
way, so each site presents different displacement rates in envrionmental
spaces and as a consequence, different demographic rates. So we can
establish that for each site \(i\) in geographic space \(\mathbf{G}\),
there is a population size \(N_i\) that changes with respect to a
fitness function dependent on time-varying environmental conditions:

\[
N_i(\vec{e}_i(t)) \propto  S_i(\vec{e}_i(t))
\]

\hypertarget{genetic-variation-and-effective-population-size}{%
\paragraph{Genetic variation and effective population
size}\label{genetic-variation-and-effective-population-size}}

On the other hand, loss of heterozygosity may be related to the amount
of genetic variation present in the absence of mutation and selection,
so theoretically, one would expect a strong correlation between
effective population size and heterozygosity on the basis of population
genetics. We start from the assumption that in the absence of gene flow
between subpopulations, the rate of fixation by isolation depends on the
effective population size \(N_e\) when subpopulations diverge from a
common ancestor in \(t\) generations. Wright (1943):

\[
F_{st}(t) = 1 - e^{-kt/N_e } 
\]

where we assume, according to Nei (Nei 1986), that genetic drift within
each subpopulation causes the average heterozygosity between populations
(\(H_s\)) to approach zero (\(F_{st} = 1 - H_s/H_t\)), i.e., for \(t=0\)
\(F_{st} = 0\) and for very large \(t\) the index is \(1\). that is, for
\(t=0\) \(F_{st} = 0\) and for very large \(t\) the index is \(1\). We
have included a constant \(k\) without loss of generalization.

Using
\(e(x) = \sum_{k=0}^{\infty} \frac{x^k}{k!} = 1 + x + \frac{x^2}{2} + ...\)
and leaving to first order for \(N_e\) large enough

\[
F_{st} \approx \frac{kt}{N_e} + C
\]

Moreover, the effective population can be estimated as the harmonic mean
of the population in a demographic time series (Karlin 1968) for
\(\tau\) generations
\(\frac{1}{N_e} \approx \frac{1}{\tau}\Big({\frac{1}{N_1} + \frac{1}{N_2} + ... + \frac{1}{N_\tau} } \Big)\)
for each \(i\) site in \(\mathbf{G}\). We substitute the harmonic mean
into the above equation, considering that although the number of
generations is not exactly equal to the time of the fixation index, they
are expected to be very close:

\[
F_{i}^{st} \approx \Big (kt\Big)  \frac{1}{\tau}\sum_{j=1}^{\tau}\frac{1}{N_{ij}} 
\]

\hypertarget{relationship-between-the-fixation-index-and-environmental-suitability}{%
\paragraph{Relationship between the fixation index and environmental
suitability}\label{relationship-between-the-fixation-index-and-environmental-suitability}}

We can infer locally for each population \(N_i\) in the geography expect
to be proportional to the environmental suitability for each time
considered, For simplification in the index we select a \(N_i\)
population in the geography and infer that locally, this population is
proportional to the environmental suitability in each time considered,
i. e. \(N_{j} \propto S_{j}(\vec{e}_{j})\) (each index \(j\) is for the
generation time), as we describe above, so we can asume an harmonic mean
of suitabilities
\(\frac{1}{\tau}\sum_{j = 1}^{\tau} \frac{1}{S_j} = \frac{1}{S}\) and
substitue in the previous equation:

\[
F_{st} \approx \Big (kt\Big)  \frac{1}{\tau}\sum_{j=1}^{\tau}\frac{1}{S_j(\vec{e_j})}
\]

We call this harmonic mean of environmental suitability
\emph{Suitability prevalence index }(\(S\)) given between \(0\) and
\(1\), since it reflects the rate of population growth as a function of
environment over a period of time. Thus we have a way to check that the
loss of heterozygosity (through the fixation index \(F_{st}\)) is
inversely proportional to the suitability prevalence for an \(i\) site
in the geography:

\[
F_{i}^{st}= \frac{\beta_0}{S_i} + \beta_1
\] If we consider all sites in the available geography, we call this
harmonic \emph{Suitability Prevalence Area}
\(\mathbf{S} = \{S_i | S_i > k\}\) since we expect that areas where the
\(F_{st}\) is lower are in areas where the prevalence over time is
higher, i.e.~environmental conditions remain suitable for a significant
period of time and therefore no effects of genetic drift of populations
have occurred. Conversely, it is expected that in populations with a
high fixation index value, fitness will not have prevailed over the same
time interval even though conditions currently exist and populations are
present.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

We summarize our 5 steps predictive approach in the figure 2.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth,height=\textheight]{all_figures/figure_2.png}
\caption{Realized niche in environmental space. A point dynamically
moves along different time slices}
\end{figure}

We propose a method to find the Suitability Prevalence Area (SPA) as an
index with a dual purpose: 1) to find endemic areas to delimit the
historical distribution of species and 2) to explain patterns of genetic
diversity.

The method for estimating SPA consists of a series of 4 steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Model the potential distribution of a taxon with present-day
  environmental conditions and obtain an environmental suitability map.
\item
  Project the potential distribution model to various environmental
  scenarios in the past with constant time steps and obtain the
  environmental suitability for each time scenario.
\item
  Characterize past environments and retain projections only for sites
  with analogous environmental conditions to the present.
\item
  Calculate the harmonic mean of the suitability for all the sites all
  time slices.
\end{enumerate}

Subsequently, one application of the SPA is to explain genetic diversity
(the fixation index) as a function of the SPA and project it to the
geography. For this it is necessary to perform a series of steps, which
consist of:

\begin{itemize}
\tightlist
\item
  Obtain georeferenced points of genetically structured populations with
  fixation index values.
\item
  Extract values from the SPA map with these points.
\item
  A linear model of the fixation index value with respect to the SPA is
  performed.
\item
  With the linear model all SPA values of suitability are interpolated.
\item
  A map is generated with all the values interpolated by the model.
\end{itemize}

To test our model we use the Aberti squirrel (\emph{Sciurus aberti}) as
a case study because:

\begin{itemize}
\tightlist
\item
  Since it is a medium-sized mammal (0.5 kg) its identification is easy.
\item
  The subspecies are geographically isolated, since they are found in
  pine forests surrounded by desert. Therefore we can consider the
  assumption of isolation by distance (Wright 1943).
\item
  There are sufficient occurrence records for the entire range of the
  known distribution.
\item
  It has low speciation rate (0.25) (Upham, Esselstyn, and Jetz 2019),
  therefore, we assume that its fundamental niche is stable over time.
\item
  There is fixation index information (Bono et al. 2018) for different
  populations of the subspecies, which is useful for using SPA as a way
  to explain genetic differentiation between populations.
\end{itemize}

\hypertarget{unique-occurrence-records}{%
\subsubsection{Unique occurrence
records}\label{unique-occurrence-records}}

We have downloaded geographic information on Sciurus aberti from the
open access platform Global Biodiversity Information Facility (GBIF).
Subsequently, we eliminated duplicate records, those without precise
coordinates and those with uncorroborated collection information. Once
the unique occurrence records were obtained, we filtered the geographic
space in order to avoid overpredictions in the spatial distribution
models caused by the agglomeration of occurrence points. The filtering
consisted of superimposing the records in the geographic space, then we
made 5 latitudinal slices of 5 degrees of separation between each one,
from latitude 20 degrees south to latitude 45 degrees north, taking as
criteria the known distribution of the species and the ecoregions of
North America, level 1 (downloaded from the US Environmental Protection
Agency, 2010). Once the geographic space was segmented, we left the same
number of records for each band (30 records per band) obtaining a total
of 150 unique records of presence.

\hypertarget{environmental-characterization}{%
\subsubsection{Environmental
characterization}\label{environmental-characterization}}

The environmental information to generate the models, from the present
and the reconstruction to the past, was obtained from Pastclim 1.2, an R
statistical software package designed to download and manipulate
paleoclimatic datasets. In our particular case we chose the Beyer at al.
Beyer, Krapp, and Manica (2020) set as it has environmental information
available up to 120 000 BE, significantly evolutionary time, at 2000
year time intervals based on Global Circulation Models HadCM3
(Singarayer and Valdes 2010) and HadAM3H (Valdes et al. 2017). The
environmental dataset has 17 bias-corrected bioclimatic variables,
reduced spatial scale and spatial resolution of 0.5°square cells.

\hypertarget{accessibility-region-textbfm.}{%
\subsubsection{\texorpdfstring{Accessibility region
(\(\textbf{M}\)).}{Accessibility region (\textbackslash textbf\{M\}).}}\label{accessibility-region-textbfm.}}

Species distribution models according to the BAM diagram (Soberon and
Peterson 2005) states that the geographic range occupied by a species
(\(\textbf{G}_o\)) is the region of appropriate assemblages in terms of
abiotic conditions. The region \(\textbf{M}\) represents in geography
areas in which the species has access due to its colonization
capabilities and the structure of geographic barriers within a specific
period of time (Soberón and Nakamura 2009).

In this study we selected a set of North American ecoregions for
modeling to delimit our Accessibility region. We took into account
ecoregions that overlap with the known distribution of the species and
also took into account neighboring ecoregions to increase the potential
accessibility in projection scenarios to the past. Thus, regions 5, 6,
9, 10, 12, 13 of the level I ecoregions (\url{https://www.epa.gov/})
were selected.

\hypertarget{niche-models}{%
\subsubsection{Niche models}\label{niche-models}}

\hypertarget{current-distribution}{%
\paragraph{Current distribution}\label{current-distribution}}

Estimating niche boundaries for species occurrence is called ecological
niche modeling (ENM), and when the emphasis is on geographic
distribution, it is known as species distribution modeling (SDM) (Guisan
and Thuiller 2005; Peterson and Soberón 2012; Saupe et al. 2012). We
model the current distribution of S. aberti using the MAXENT 3.4.4
algorithm (Phillips, Anderson, and Schapire 2006), which is a
correlative model based on the maximum entropy principle used to
estimate species distributions. The output of Maxent is a relative
occurrence rate (ROR) interpreted as a probability of habitat
suitability given observed environmental conditions by presence-only
points (Merow, Smith, and Silander Jr 2013).

The parameterization was performed with the default values of the
program with the exception of the ``extrapolate'' and ``clamping''
options to avoid artificial extrapolations in the extreme values of the
climatic variables used in the models. To calibrate the models we used
70\% of the records and the remaining 30\% to validate them.

In order to give statistical certainty and to consider uncertainty, as
well as to reduce overfitting and to have a suitable model to
extrapolate to past scenarios, 10 replications were performed for each
model using a cross-validation with the training data. In the end, the
average model of these replications was considered as the result.

We evaluated the statistical significance of the resulting models for
the present using the partial ROC test (Peterson, Papeş, and Soberón
2008), which is a modification of the ROC (receiver Operation
Characteristic) test. The results of this test are proportions (ratios)
of the area under the curve of the model with respect to a null model,
product of repetitions that allow to statistically evaluate the areas
under the curve (AUC) in relation to that expected by chance (Peterson,
Papeş, and Soberón 2008), where a value derived by chance would be \(0\)
and an acceptable value, according to the proportion of minimum omission
errors tolerated in the model, would be greater than \(1\). A different
set of test data was used on the average model of the replicates and in
this way the reduction of overfitting was guaranteed.

\hypertarget{reconstruction-to-the-past}{%
\paragraph{Reconstruction to the
past}\label{reconstruction-to-the-past}}

The average model obtained from Maxent was extrapolated to environmental
conditions into the past. Environmental layers were taken from pastclim
(Leonardi et al. 2023) in 2000 year time slices to cover a total of 120
000 years. For each scenario into the past, 10 replicates were
cross-validated and averaged.

\hypertarget{analogy-of-environmental-conditions}{%
\paragraph{Analogy of environmental
conditions}\label{analogy-of-environmental-conditions}}

To be certain that areas in the past are climatically analogous to the
present we implemented the mobility-orient parity (MOP) test. This
method generates a value from \(0\) to \(1\) according to the
environmental analogy of past conditions to the present to eliminate
geographic areas in the projections to the past outside of climates
analogous to the present (Owens et al. 2013).

\hypertarget{suitability-prevalence-area-spa}{%
\subsubsection{Suitability Prevalence Area
(SPA)}\label{suitability-prevalence-area-spa}}

Finally, to generate the SPA, both the average projection of the present
and all the average projections of the past are considered. The output
files are read in raster format.

A harmonic mean is performed as follows: For each layer the inverse of
each cell value is obtained. Then the inverse values are summed by cells
and divided by the number of scenarios. Finally, the inverse of each
cell is obtained. It is important to consider that the Maxent outputs
may contain very small values close to zero in many cells, for this
reason we added a small amount of \(0.001\) to all the values in all the
maps so that when obtaining the harmonic mean we could have an inverse
value for values close to zero without altering the biological sense of
environmental suitability.

\hypertarget{endemic-historic-area}{%
\subsubsection{Endemic historic Area}\label{endemic-historic-area}}

To calculate the historic endemic area, the SPA result is taken and a
threshold is applied to generate a binary map according to the
\(\mathbf{S}\) equation. This threshold is the proportion of time that
pixel has had ideal conditions. For our study, we consider the value of
90\% since for a period of 120 000 years it is a coverage of 100 000
years, which is a significantly evolutionary time to fix genetic
patterns at the population level.

\hypertarget{fixation-index-as-a-functino-of-spa}{%
\subsubsection{Fixation index as a functino of
SPA}\label{fixation-index-as-a-functino-of-spa}}

To test our hypothesis of genetic diversity as a function of
environmental suitability prevalence, from a set of genetically
structured populations, a matrix of fixation indices between populations
is considered. For this work we considerer the data from @Bono et al.
(2018). It consideres 10 populations along the known distribution and
provides values of \(F_{st}\) between this populations in matrix form.
The matrix was transformed into a table as pairs between populations
with index values between each pair. The values of the population with
itself, i.e.~\(0\), were included. Subsequently, georeferences were
obtained for each population considered in the fixation index matrix.
From this geo-referencing the SPA value for each population was
extracted. Finally, the SPA value was aligned in the table of fixation
index pairs. In total the table has 4 columns: source population, target
population, fixation index and SPA of the target population. This table
was separated by source population groups to generate a set of linear
models between SPA values and fixation indexes \(F_{st}\). An analysis
was performed for each source population with the hypothesis that the
correlation between SPA and fixation index exists and is non-zero if a
dispersal process has occurred from the historical endemic area of the
taxon.

\hypertarget{fixation-index-projected-in-geography}{%
\subsubsection{Fixation index projected in
geography}\label{fixation-index-projected-in-geography}}

The linear model that presents the highest value of \(R^2\) is
considered suitable to make a geographic projection of the fixation
index and observe genetic diversity in geographic space. To generate
this map, the SPA values are taken and a table of coordinates and SPA
values is generated for each cell of the raster. The SPA values are then
extrapolated to obtain fixation index values from the linear model
obtained previously. This column is added to the coordinate table and a
raster is generated with the projected \(F_{st}\) values. Finally, a
cutoff of this raster is performed using a map of the distribution of
the species at present with a cutoff threshold of 90\% in binary format
(where \(1 =\)presence and \(0 =\) absence), since the records of the
presence localities were obtained from bases that might contain some
errors in the data (Phillips, Anderson, and Schapire 2006).

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{current-and-past-distribution}{%
\subsubsection{Current and past
distribution}\label{current-and-past-distribution}}

We show the species distribution model for \emph{Sciurus aberti} along
the time line of the projected area per time slice. We show 5 notable
moments in time line. and 5 notable points

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{all_figures/figure_3.png}
\caption{A) Species distribution model for \emph{Sciurus aberti} present
distribution. B) Relative projected area for SDM in pixels acoording a
threshold of 0.9 suitability for each past time-sliced environmental
conditions. We can observe a maximum area corresponding to the last
interglacial era and an inverted U-shape in the trend of the area along
the time considered. Also It can be noted that the actual distribution
has similar area to the coditions 120 000 years ago. B) and D)
corresponds to distribution peaks along the time scale and C) and E)
corresponds to minimum}
\end{figure}

\hypertarget{spa-and-endemic-historic-area}{%
\subsubsection{SPA and Endemic Historic
Area}\label{spa-and-endemic-historic-area}}

We get the suitability presence area (figure 2) delmited by a 0.5 and
0.9 threshold (Figure 4). The endemic region where the enviornmental
conditions 90 \% del tiempo is the Sierra Madre Occidental, in the
limits of Sonora, Chihuahua, Sinaloa and Durango. We correspond this
Area to the \emph{Sciurus aberti barbieri} subspecies. Also is notable
that at 0.5 threshold, other regions seems to have stable conditions. We
observe the area close to Coronado National Forest as well as Gila
National Forest and Mescalero Reservation as suitable areas along time
in the United States and some Areas of the Volcanic Belt in Mexico
(Puebla, north of Guerrero State and Jalisco.\\

\begin{figure}
\centering
\includegraphics{all_figures/figure_4.png}
\caption{Suitability prevalence area. Here we observe the harmonic mean
of suitability for each pixel along the distribucion from 0 to 1. The
bold lines shows the the area when the harmonic mean es above 0.5
threshold (blud line) and 0.9 threshold (black line)}
\end{figure}

\hypertarget{fst-as-a-funciton-of-spa-and-fixation-index-projected-in-geography}{%
\subsubsection{Fst as a funciton of SPA and Fixation index projected in
geography}\label{fst-as-a-funciton-of-spa-and-fixation-index-projected-in-geography}}

For all the linear regressions performed (Table 1), the only one that
had statistical significance is the one for the Sciurus aberti barberi
population with an R-squared value of 0.794.

\begin{longtable}[]{@{}ccc@{}}
\caption{R squared from linear models (fixation index as a function of
suitability prevalence area). The only R squared with statistical
significance is barberi}\tabularnewline
\toprule\noalign{}
population & r.squared & p.value \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
population & r.squared & p.value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
S. a. aberti Carson-SFW & 0.000 & 0.962 \\
S. a. aberti Coconino-Gila & 0.131 & 0.304 \\
S. a. aberti MT-Zuni & 0.002 & 0.894 \\
S. a. aberti San Juan & 0.000 & 0.952 \\
S. a. barberi & 0.794 & 0.001 \\
S. a. chuscensis E & 0.002 & 0.892 \\
S. a. chuscensis W & 0.018 & 0.713 \\
S. a. ferreus Carson E & 0.040 & 0.581 \\
S. a. ferreus Pike & 0.000 & 0.967 \\
S. a. kaibabensis & 0.005 & 0.841 \\
\end{longtable}

It can be noted that S. a. aberti Coconino-Gila had an R-squared value
greater than 0 (0.131) but it is not statistically significant.
Furthermore, in our approach we imputed the value of the origin with an
Fst of zero. To show that the model hypothesis still holds, we performed
a regression for barberi without the point of origin imposed (Figure
5b), in addition to eliminating the S. aberti chuscensis populations due
to their low uncertainty in georeference, obtaining an R-squared of 0.72
(p value = 0.016).

\begin{figure}
\includegraphics[width=0.5\linewidth]{all_figures/figure_5a} \includegraphics[width=0.5\linewidth]{all_figures/figure_5b} \caption{Linear regression for S a. barberi population A) SPA in all populations B) withouth barberi and chusensis populations. We can observe that in both cases the fixation index is explained with the SPA and the hypothesis holds without the imputation of the origin of the population (this means to add a 0 $F_{st}$ value for barberi population)}\label{fig:figures-side}
\end{figure}

From the model we extrapolated the values from SPA along the geography
to new \(F_{st}\) values to get a geographical projection of fixation
index from \emph{S. a. barberi} population. (Figure 6)

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{all_figures/figure_6.png}
\caption{Projection of the \(F_{st}\) model to the geography in the
actual distribution of \emph{S aberti}}
\end{figure}

\hypertarget{references}{%
\subsubsection*{References}\label{references}}
\addcontentsline{toc}{subsubsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-beyer2020high}{}}%
Beyer, Robert M, Mario Krapp, and Andrea Manica. 2020.
{``High-Resolution Terrestrial Climate, Bioclimate and Vegetation for
the Last 120,000 Years.''} \emph{Scientific Data} 7 (1): 236.

\leavevmode\vadjust pre{\hypertarget{ref-bono2018genome}{}}%
Bono, Jeremy M, Helen K Pigage, Peter J Wettstein, Stephanie A Prosser,
and Jon C Pigage. 2018. {``Genome-Wide Markers Reveal a Complex
Evolutionary History Involving Divergence and Introgression in the
Abert's Squirrel (Sciurus Aberti) Species Group.''} \emph{BMC
Evolutionary Biology} 18 (1): 1--17.

\leavevmode\vadjust pre{\hypertarget{ref-burgin2018many}{}}%
Burgin, Connor J, Jocelyn P Colella, Philip L Kahn, and Nathan S Upham.
2018. {``How Many Species of Mammals Are There?''} \emph{Journal of
Mammalogy}. Oxford University Press US.

\leavevmode\vadjust pre{\hypertarget{ref-guisan2005predicting}{}}%
Guisan, Antoine, and Wilfried Thuiller. 2005. {``Predicting Species
Distribution: Offering More Than Simple Habitat Models.''} \emph{Ecology
Letters} 8 (9): 993--1009.

\leavevmode\vadjust pre{\hypertarget{ref-hedrick2005standardized}{}}%
Hedrick, Philip W. 2005. {``A Standardized Genetic Differentiation
Measure.''} \emph{Evolution} 59 (8): 1633--38.

\leavevmode\vadjust pre{\hypertarget{ref-karlin1968rates}{}}%
Karlin, Samuel. 1968. {``Rates of Approach to Homozygosity for Finite
Stochastic Models with Variable Population Size.''} \emph{The American
Naturalist} 102 (927): 443--55.

\leavevmode\vadjust pre{\hypertarget{ref-krapp2021statistics}{}}%
Krapp, Mario, Robert M Beyer, Stephen L Edmundson, Paul J Valdes, and
Andrea Manica. 2021. {``A Statistics-Based Reconstruction of
High-Resolution Global Terrestrial Climate for the Last 800,000
Years.''} \emph{Scientific Data} 8 (1): 228.

\leavevmode\vadjust pre{\hypertarget{ref-leonardi2023pastclim}{}}%
Leonardi, Michela, Emily Y Hallett, Robert Beyer, Mario Krapp, and
Andrea Manica. 2023. {``Pastclim 1.2: An r Package to Easily Access and
Use Paleoclimatic Reconstructions.''} \emph{Ecography}, e06481.

\leavevmode\vadjust pre{\hypertarget{ref-lira2014relationship}{}}%
Lira-Noriega, Andrés, and Joseph D Manthey. 2014. {``Relationship of
Genetic Diversity and Niche Centrality: A Survey and Analysis.''}
\emph{Evolution} 68 (4): 1082--93.

\leavevmode\vadjust pre{\hypertarget{ref-merow2013practical}{}}%
Merow, Cory, Matthew J Smith, and John A Silander Jr. 2013. {``A
Practical Guide to MaxEnt for Modeling Species' Distributions: What It
Does, and Why Inputs and Settings Matter.''} \emph{Ecography} 36 (10):
1058--69.

\leavevmode\vadjust pre{\hypertarget{ref-nei1986definition}{}}%
Nei, Masatoshi. 1986. {``Definition and Estimation of Fixation
Indices.''} \emph{Evolution} 40 (3): 643--45.

\leavevmode\vadjust pre{\hypertarget{ref-nogues2009predicting}{}}%
Nogués-Bravo, David. 2009. {``Predicting the Past Distribution of
Species Climatic Niches.''} \emph{Global Ecology and Biogeography} 18
(5): 521--31.

\leavevmode\vadjust pre{\hypertarget{ref-owens2013constraints}{}}%
Owens, Hannah L, Lindsay P Campbell, L Lynnette Dornak, Erin E Saupe,
Narayani Barve, Jorge Soberón, Kate Ingenloff, et al. 2013.
{``Constraints on Interpretation of Ecological Niche Models by Limited
Environmental Ranges on Calibration Areas.''} \emph{Ecological
Modelling} 263: 10--18.

\leavevmode\vadjust pre{\hypertarget{ref-peterson2008rethinking}{}}%
Peterson, A Townsend, Monica Papeş, and Jorge Soberón. 2008.
{``Rethinking Receiver Operating Characteristic Analysis Applications in
Ecological Niche Modeling.''} \emph{Ecological Modelling} 213 (1):
63--72.

\leavevmode\vadjust pre{\hypertarget{ref-peterson2012species}{}}%
Peterson, A Townsend, and Jorge Soberón. 2012. {``Species Distribution
Modeling and Ecological Niche Modeling: Getting the Concepts Right.''}
\emph{Natureza \& Conserva{ç}{ã}o} 10 (2): 102--7.

\leavevmode\vadjust pre{\hypertarget{ref-phillips2006maximum}{}}%
Phillips, Steven J, Robert P Anderson, and Robert E Schapire. 2006.
{``Maximum Entropy Modeling of Species Geographic Distributions.''}
\emph{Ecological Modelling} 190 (3-4): 231--59.

\leavevmode\vadjust pre{\hypertarget{ref-saupe2012variation}{}}%
Saupe, EE, V Barve, CE Myers, J Soberón, N Barve, CM Hensz, AT Peterson,
H Lc Owens, and A Lira-Noriega. 2012. {``Variation in Niche and
Distribution Model Performance: The Need for a Priori Assessment of Key
Causal Factors.''} \emph{Ecological Modelling} 237: 11--22.

\leavevmode\vadjust pre{\hypertarget{ref-singarayer2010high}{}}%
Singarayer, Joy S, and Paul J Valdes. 2010. {``High-Latitude Climate
Sensitivity to Ice-Sheet Forcing over the Last 120 Kyr.''}
\emph{Quaternary Science Reviews} 29 (1-2): 43--55.

\leavevmode\vadjust pre{\hypertarget{ref-soberon2005interpretation}{}}%
Soberon, Jorge, and A Townsend Peterson. 2005. {``Interpretation of
Models of Fundamental Ecological Niches and Species' Distributional
Areas.''}

\leavevmode\vadjust pre{\hypertarget{ref-soberon2009niches}{}}%
Soberón, Jorge, and Miguel Nakamura. 2009. {``Niches and Distributional
Areas: Concepts, Methods, and Assumptions.''} \emph{Proceedings of the
National Academy of Sciences} 106 (supplement\_2): 19644--50.

\leavevmode\vadjust pre{\hypertarget{ref-soberon2020shape}{}}%
Soberón, Jorge, and A Townsend Peterson. 2020. {``What Is the Shape of
the Fundamental Grinnellian Niche?''} \emph{Theoretical Ecology} 13 (1):
105--15.

\leavevmode\vadjust pre{\hypertarget{ref-thorup2021response}{}}%
Thorup, Kasper, Lykke Pedersen, Rute R Da Fonseca, Babak Naimi, David
Nogués-Bravo, Mario Krapp, Andrea Manica, et al. 2021. {``Response of an
Afro-Palearctic Bird Migrant to Glaciation Cycles.''} \emph{Proceedings
of the National Academy of Sciences} 118 (52): e2023836118.

\leavevmode\vadjust pre{\hypertarget{ref-upham2019inferring}{}}%
Upham, Nathan S, Jacob A Esselstyn, and Walter Jetz. 2019. {``Inferring
the Mammal Tree: Species-Level Sets of Phylogenies for Questions in
Ecology, Evolution, and Conservation.''} \emph{PLoS Biology} 17 (12):
e3000494.

\leavevmode\vadjust pre{\hypertarget{ref-valdes2017bridge}{}}%
Valdes, Paul J, Edward Armstrong, Marcus PS Badger, Catherine D
Bradshaw, Fran Bragg, Michel Crucifix, Taraka Davies-Barnard, et al.
2017. {``The BRIDGE HadCM3 Family of Climate Models: HadCM3@ Bristol V1.
0.''} \emph{Geoscientific Model Development} 10 (10): 3715--43.

\leavevmode\vadjust pre{\hypertarget{ref-wright1931evolution}{}}%
Wright, Sewall. 1931. {``Evolution in Mendelian Populations.''}
\emph{Genetics} 16 (2): 97.

\leavevmode\vadjust pre{\hypertarget{ref-wright1943isolation}{}}%
---------. 1943. {``Isolation by Distance.''} \emph{Genetics} 28 (2):
114.

\end{CSLReferences}

\end{document}