### 2023-07-29 abstract.Rmd
changes in manuscript

---
title: Suitability prevalence area index in late quaternary explains genetic diversity in Tassel eared Squirrels
author: "Norma Hernandez, Angel Robles Nathan Upham"
date: "2023-06-27"
includes:
  in_header: preamble.tex
output: 
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex 
    keep_tex: true
bibliography: R21_citation.bib
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

The current distributions of species do not always correspond to their
historical distributions over evolutionarily significant periods. This
is because environmental conditions are not static over time; species
tend to distribute where conditions are most favorable.

It follows from this relationship between environment and species
distribution that population size also varies with time. This variation in population size is related to the effective population size $N_e$, so that an index reflecting changes in environmental
conditions in geography might be expected to be related to $N_e$ and
thus to indicators of population structure such as the fixation index
$F_{st}$. Thus, it is possible to relate patterns of changes in the
distribution of species to the genetic structure of their populations
using a statistical model that explains this relationship.

With this approach we can predict the geographic pattern of population
structure from environmental information. This approach is strongly
driven by advances in currently available climate simulations
[@leonardi2023pastclim; @krapp2021statistics], as well as next
generation sequencing data, and supported by both ecological niche
[@thorup2021response; @nogues2009predicting] and population genetics
theories [@lira2014relationship].

In this work we propose a method to find the Suitability Prevalence Area (SPA) as an index with a double purpose: 1) to find endemic areas to delimit the historical distribution of the species and 2) to explain the patterns of genetic diversity.

To obtain the SPA, we performed a historical reconstruction of the
geographical range back to 120 000 BC at 2 000 year intervals and
recorded the environmental suitability at each site.

Subsequently, we delimited historical endemic areas to locations where
the prevalence of suitability remained at 90\% during this period. From the fixation index calculated with respect to populations in the
historical endemic areas, a statistical model of the fixation index as a function of SPA was performed. With this statistical model, the fixation index values were projected to the current distribution to obtain a map with the geographic pattern of this index.

As a case study we consider squirrels (*Sciurus aberti*), a species
currently distributed in disjunct patches from the southern Rocky
Mountains in the United States to the northern Sierra Madre Occidental
in Mexico for which it is possible to find reliable information on both the genetic structure of its populations and its current distribution [@bono2018genome; @burgin2018many].

Our results reveal that suitability prevalence corresponds to the
fixation index of *S. aberti* populations with respect to a source
population. Populations closer to the historical endemic area present a higher genetic diversity and a lower $F_{st}$ value. Finally, this study allows us to add a biogeographic explanation to the results obtained with population genetic methods and to generate maps of this structure as tools to support conservation with a perspective that integrates both population genetics and historical patterns of species distribution.